Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies

### 2025-02-27 paper.tex
added some initial results/obs

\newcommand{\figbig}[0]{0.4\textwidth}
\newcommand{\figmed}[0]{0.4\textwidth}
\newcommand{\figsmall}[0]{0.3\textwidth}

@@

In this work, we apply a light curve attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station. Given an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values. Specifically, we estimate a full distribution for each candidate light curve and compute the likelihood of the observations being a sample from that distribution. To solve the optimization problem, we create a number of sampled candidate solutions, deterministically arranged throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm. This way, our optimization procedure reliably converges to all maximum likelihood estimates of the attitude state. Low-error minima are then clustered to identify families of nearby solutions with similar likelihood. The final output of our method is an array of the best solutions, ranked by their likelihood. Formulating the problem this way inherently accounts for the multiple classes of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, as well as the noise on the light curve.

@@

In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance.

@@

@@

[\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]

\bibliography{cache/refs.bib}

% \begin{thebibliography}{}

% \bibitem{author73}
% Author T., (1973). Astrophysical Quantities, Athlone Press

% \bibitem{nobody97}
% Nobody B., Somebody G., Who D., et~al., (1997). {\it The book}, Publisher, ed. 2

% \bibitem{smith96}
% Smith A., Jones B., (1996). The new discovery, {\it Other Journal}, {\bf 223}(1), 1029--1101

% \end{thebibliography}
\end{document}

>>>>

\newcommand{\figbig}[0]{0.5\textwidth}
\newcommand{\figmed}[0]{0.4\textwidth}
\newcommand{\figsmall}[0]{0.3\textwidth}

@@

In this work, we apply a light curve attitude inversion algorithm to real observations taken by the Purdue Optical Ground Station. Given an estimate of the shape and reflective properties of the object, our procedure searches the space of initial orientations, angular velocities, and inertia tensors to find initial conditions that produce light curves with low error compared to the observed values. Specifically, we estimate a full distribution for each candidate light curve and compute the likelihood of the observations being a sample from that distribution. To solve the optimization problem, we create a number of sampled candidate solutions, scattered throughout the solution space, and perform nonlinear optimization on each using the BFGS algorithm. This way, our optimization procedure reliably converges to all maximum likelihood estimates of the attitude state. Low-error minima are then clustered to identify families of nearby solutions with similar likelihood. The final output of our method is an array of the best solutions, ranked by their likelihood. Formulating the problem this way inherently accounts for the multiple classes of ambiguities resulting from the object and observation geometry, the physical constraints of torque-free rigid body motion, as well as the noise on the light curve.

@@

In this work, the light curve and its standard deviation are always assumed to be expressed in a linear unit, e.g., analog-to-digital units (ADU) or irradiance proportional to $W/m^2$.

@@

\section{Real Data Considerations}

\section{Results}

The two objects selected for study in this work are a Delta I upper stage and the ECHOSTAR 2 satellite.

\subsection{Delta I Upper Stage}

The Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}. It is $d=0.93$ meters in diameter and is approximately $h=1.7$ meters in length \cite{star37e_astronautix, star37_gunter}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figsmall]{star37e.JPG}
  \caption{The Star-37E upper stage \cite{star37_af}.}
  \label{fig:star37e}
\end{figure}

\textit{Computing its MOI}

We compute the inertia tensor of the Star-37E by approximating its geometry as a hollow cylinder of the same aspect ratio such that \cite{serway2019}:

\begin{align}
  J_a &= \frac{1}{2} m \left(r_o^2+r_i^2\right) \\
  J_t &= \frac{1}{12} m \left(3 \left(r_o^2+r_i^2\right) + h^2\right). \\
\end{align}

Here, we choose the inner radius to be $r_i=0.45*d$, the outer radius to be $r_o=0.5d$, and $m=1$ kilogram -- dividing each moment of inertia by a constant does not change the dynamics -- yielding $J_a = 15.3 \: [\text{kg} \cdot m^2]$ and $J_t = 27.9 \: [\text{kg} \cdot m^2]$.

[\textit{MODEL PIC 1}]

\subsection{ECHOSTAR II}

ECHOSTAR II (COSPAR ID 1996-055A) was a communications satellite on the AS-7000 bus \cite{as7000_astronautix}, shown in Figure \ref{fig:echostar1}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figbig]{EchoStar-1_image.jpg}
  \caption{ECHOSTAR II artist's rendition \cite{as7000_astronautix}.}
  \label{fig:echostar1}
\end{figure}

[\textit{MODEL PIC 2}]

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figbig]{sphx_glr_plot_models_001.png}
  \caption{Simplified ECHOSTAR II box-wing model used for inversion.}
  \label{fig:echostar1_simple}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figbig]{sphx_glr_plot_models_002.png}
  \caption{Simplified Star-37E model used for inversion.}
  \label{fig:star37e_simple}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_001_2_00x.png}
  \caption{Selected light curve for the Delta (Star-37E) rocket body, observed by the Purdue Optical Ground Station}
  \label{fig:rb_lc_obs}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\figbig]{sphx_glr_plot_lcs_002_2_00x.png}
  \caption{Selected light curve for ECHOSTAR II, observed by the Purdue Optical Ground Station}
  \label{fig:sat_lc_obs}
\end{figure}

@@

% [\textit{TODO: Acknowledge AIUB if we end up using some of their data for a GLONASS light curve}]

\bibliography{cache/refs.bib}
\end{document}
